% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  fontsize=13pt,
  english,
  a4paper,
  openany, a4paper, oneside]{book}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Open Collaboration Guide},
  pdfauthor={Daniel Antal},
  pdflang={en},
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=3cm, right=3cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{amsthm}
\makeatletter
\def\thm@space@setup{%
  \thm@preskip=8pt plus 2pt minus 4pt
  \thm@postskip=\thm@preskip
}
\makeatother
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{apalike}

\title{Open Collaboration Guide}
\author{Daniel Antal}
\date{2021-04-16}

\begin{document}
\maketitle

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{1}
\tableofcontents
}
\setstretch{1.1}
\hypertarget{welcome}{%
\chapter*{Welcome}\label{welcome}}
\addcontentsline{toc}{chapter}{Welcome}

Placeholder

\hypertarget{intro}{%
\chapter{Introduction}\label{intro}}

Placeholder

\hypertarget{definitions}{%
\section{Definitions}\label{definitions}}

\hypertarget{pledge}{%
\section{Code of Conduct}\label{pledge}}

\hypertarget{collaboration-tools}{%
\section{Collaboration Tools}\label{collaboration-tools}}

\hypertarget{keybase}{%
\subsection{Instant messaging: Keybase}\label{keybase}}

\hypertarget{github}{%
\subsection{Git \& Github}\label{github}}

\hypertarget{rstudio-ide-other-ide}{%
\subsection{Rstudio IDE \& other IDE}\label{rstudio-ide-other-ide}}

\hypertarget{simple-intro}{%
\section{Simple Introduction}\label{simple-intro}}

\hypertarget{markdown}{%
\subsection{Markdown}\label{markdown}}

\hypertarget{inspiration}{%
\section{Inspiration \& Recommended Reading}\label{inspiration}}

\hypertarget{books}{%
\subsection{Books}\label{books}}

\hypertarget{why-weapons-of-math-destruction}{%
\subsubsection{Why: Weapons of Math Destruction}\label{why-weapons-of-math-destruction}}

\hypertarget{how-metadata}{%
\subsubsection{How: Metadata}\label{how-metadata}}

\hypertarget{critically-data-feminism}{%
\subsubsection{Critically: Data Feminism}\label{critically-data-feminism}}

\hypertarget{blogposts}{%
\subsection{Blog posts \& Podcasts}\label{blogposts}}

\hypertarget{data-curation}{%
\chapter{Data Curation}\label{data-curation}}

Placeholder

\hypertarget{eu-datathlon}{%
\section{EU Datathlon 2021}\label{eu-datathlon}}

\hypertarget{observatories}{%
\section{Observatories}\label{observatories}}

\hypertarget{meta-observatory}{%
\section{Meta Observatory}\label{meta-observatory}}

\hypertarget{emo}{%
\subsection{European Music Observatory}\label{emo}}

\hypertarget{strengths}{%
\section{Strengths}\label{strengths}}

\hypertarget{acquisition}{%
\chapter{Data Acquistion}\label{acquisition}}

Whatever is the source of the data we use it, we never trust it fully. We need check its strength and weaknesses, and bring it to a complete, documented and tidy form.

\hypertarget{collect-metadata}{%
\section{Metadata}\label{collect-metadata}}

Metadata plays an important role to find whatever we acquired and use it properly. It plays an important role during the storage of data in our \protect\hyperlink{data-storage}{databases} and in \protect\hyperlink{documentation}{documentation and publication}, too. We placed it into a \protect\hyperlink{metadata}{separate chapter}.

\hypertarget{eurostat}{%
\section{Eurostat}\label{eurostat}}

\hypertarget{harmonized-survey-programs}{%
\section{Harmonized Survey Programs}\label{harmonized-survey-programs}}

\hypertarget{music-apis}{%
\section{Music APIs}\label{music-apis}}

\hypertarget{spotify-api}{%
\subsection{Spotify API}\label{spotify-api}}

We collect the following data from the Spotify API

\hypertarget{bandcamp}{%
\subsection{Bandcamp}\label{bandcamp}}

\hypertarget{data-storage}{%
\chapter{Data Storage \& Databases}\label{data-storage}}

Placeholder

\hypertarget{storage-metadata}{%
\section{Metadata}\label{storage-metadata}}

\hypertarget{dta-raw}{%
\section{Raw data assets}\label{dta-raw}}

\hypertarget{processed-individual-data}{%
\section{Processed, individual data}\label{processed-individual-data}}

\hypertarget{indicators---statistically-processed-data}{%
\section{Indicators - statistically processed data}\label{indicators---statistically-processed-data}}

\hypertarget{periodic-data-releases}{%
\section{Periodic data releases}\label{periodic-data-releases}}

\hypertarget{interactive-data-releases}{%
\section{Interactive data releases}\label{interactive-data-releases}}

\hypertarget{continous-data-releases-in-api}{%
\section{Continous data releases in API}\label{continous-data-releases-in-api}}

\hypertarget{applications}{%
\chapter{Applications}\label{applications}}

Placeholder

\hypertarget{naming-conventions}{%
\section{File, Variable Names, Value Labels}\label{naming-conventions}}

\hypertarget{path}{%
\subsection{Path}\label{path}}

\hypertarget{R-guide}{%
\subsubsection{R language}\label{R-guide}}

\hypertarget{variable-nameing-styles}{%
\subsection{Variable nameing styles}\label{variable-nameing-styles}}

\hypertarget{character-coding}{%
\subsection{Character coding}\label{character-coding}}

\hypertarget{statistical-software}{%
\section{Statistical Processing \& Indicators}\label{statistical-software}}

\hypertarget{retroharmonize}{%
\subsection{retroharmonize}\label{retroharmonize}}

\hypertarget{regions}{%
\subsection{regions}\label{regions}}

\hypertarget{iotables}{%
\subsection{iotables}\label{iotables}}

\hypertarget{machine-learning}{%
\section{Machine Learning Applications}\label{machine-learning}}

\hypertarget{listen-local}{%
\subsection{Listen Local app}\label{listen-local}}

\hypertarget{bandcamp-librarian}{%
\subsection{Bandcamp Librarian app}\label{bandcamp-librarian}}

\hypertarget{maps}{%
\section{Maps}\label{maps}}

\hypertarget{documentation}{%
\chapter{Documentation And Publications}\label{documentation}}

\hypertarget{bibliography}{%
\section{Citations, Bibliography}\label{bibliography}}

For any external literature, data source, please, store the citation information in a well-formatted, thematic \texttt{.bib} files.

We create programatically \texttt{.bib} citation information files for our data products, software releases, and published documents.

\begin{Shaded}
\begin{Highlighting}[]
\SpecialCharTok{@}\NormalTok{Manual\{R}\SpecialCharTok{{-}}\NormalTok{regions,}
\NormalTok{  title }\OtherTok{=}\NormalTok{ \{regions}\SpecialCharTok{:}\NormalTok{ Processing Regional Statistics\},}
\NormalTok{  author }\OtherTok{=}\NormalTok{ \{Daniel Antal\},}
\NormalTok{  note }\OtherTok{=}\NormalTok{ \{R package version }\DecValTok{0}\NormalTok{.}\FloatTok{1.6}\NormalTok{\},}
\NormalTok{  url }\OtherTok{=}\NormalTok{ \{https}\SpecialCharTok{:}\ErrorTok{//}\NormalTok{regions.danielantal.eu}\SpecialCharTok{/}\NormalTok{\},}
\NormalTok{  year }\OtherTok{=}\NormalTok{ \{}\DecValTok{2021}\NormalTok{\},}
\NormalTok{\}}

\SpecialCharTok{@}\NormalTok{Manual\{R}\SpecialCharTok{{-}}\NormalTok{retroharmonize,}
\NormalTok{  title }\OtherTok{=}\NormalTok{ \{retroharmonize}\SpecialCharTok{:}\NormalTok{ Ex Post Survey Data Harmonization\},}
\NormalTok{  author }\OtherTok{=}\NormalTok{ \{Daniel Antal\},}
\NormalTok{  note }\OtherTok{=}\NormalTok{ \{R package version }\DecValTok{0}\NormalTok{.}\FloatTok{1.15}\NormalTok{\},}
\NormalTok{  url }\OtherTok{=}\NormalTok{ \{https}\SpecialCharTok{:}\ErrorTok{//}\NormalTok{retroharmonize.dataobservatory.eu}\SpecialCharTok{/}\NormalTok{\},}
\NormalTok{  year }\OtherTok{=}\NormalTok{ \{}\DecValTok{2021}\NormalTok{\},}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Marking up a reference to \citep{R-retroharmonize} with \texttt{{[}@R-retroharmonize{]}} will place the citation in your selected format (here APA) to the text, and to the \protect\hyperlink{references}{references} at the end of this documentation.

\textbf{We would love to cite your work}. Please, create a well-formatted \texttt{.bib} file with your publications about the best use of data, creation of data, scientific analysis of data, etc, and save it to this repo as \texttt{name-surname.bib}.

Make sure that each \texttt{.bib} entry contains at least the \texttt{title}, \texttt{author} and \texttt{year} fields.

\hypertarget{software-releases}{%
\section{Software Releases}\label{software-releases}}

We have so far three released software products.

\begin{itemize}
\tightlist
\item
  The aim of \href{https://reprex.nl/software/retroharmonize/}{retroharmonize} is to provide tools for reproducible retrospective (ex-post) harmonization of datasets that contain variables measuring the same concepts but coded in different ways.
\item
  The \href{https://reprex.nl/software/regions/}{regions} package is an offspring of the eurostat package on \href{https://github.com/rOpenGov}{rOpenGov}. It started as a tool to validate and re-code regional Eurostat statistics, but it aims to be a general solution for all sub-national statistics. It will be developed parallel with other rOpenGov packages.
\item
  \href{https://reprex.nl/software/iotables/}{iotables} processes all the symmetric input-output tables of the EU member states, and calculates direct, indirect and induced effects, multipliers for GVA, employment, taxation. These are important inputs into policy evaluation, business forecasting, or granting/development indicator design. iotables is used by about 800 experts around the world.
\end{itemize}

They are released on CRAN, and they follow the CRAN guidelines for unit-testing and human review. As CRAN relies on extensive automated testing, the formatting standard is \emph{very strict} both for the software code and its documentation. The slightest deviation results in rejection.

\hypertarget{documentation-metadata}{%
\section{Metadata}\label{documentation-metadata}}

Metadata plays an important role in our \protect\hyperlink{data-storage}{databases} and in \protect\hyperlink{documentation}{documentation}, too. We placed it into a \protect\hyperlink{metadata}{separate chapter}.

\hypertarget{data-release}{%
\section{Data releases}\label{data-release}}

We currently work with two platforms, and we must maintain compatibility with data repositories that our clients use. Both provide some validation, version control, and give a standard \textbf{doi} to our releases.

\hypertarget{dataverse}{%
\subsection{Dataverse}\label{dataverse}}

\href{https://dataverse.org/}{dataverse.org/} has a well-supported API, and this is the choice of our first academic partner, IViR.

The Dataverse Project is being developed at Harvard's Institute for Quantitative Social Science (IQSS), along with many collaborators and contributors worldwide. The Dataverse Project was built on our experience with our earlier Virtual Data Center (VDC) project, which spanned 1997-2006 as a collaboration between the Harvard-MIT Data Center (now part of IQSS) and the Harvard University Library. Precursors to the VDC date to 1987, comprising such entities as pre-web software to automatically transfer cataloging information by FTP to other sites across campus automatically at designated times, and before that to a stand-alone software guide to local data.

\hypertarget{zenodo}{%
\subsection{Zenodo}\label{zenodo}}

\href{https://zenodo.org/}{zenodo} is the choice of the European Union, and it is likely that in the future all EU-financed research data must be published here.

The OpenAIRE project, in the vanguard of the open access and open data movements in Europe was commissioned by the EC to support their nascent Open Data policy by providing a catch-all repository for EC funded research. CERN, an OpenAIRE partner and pioneer in open source, open access and open data, provided this capability and Zenodo was launched in May 2013.

In support of its research programme CERN has developed tools for Big Data management and extended Digital Library capabilities for Open Data. Through Zenodo these Big Science tools could be effectively shared with the long-tail of research.

\hypertarget{publications}{%
\section{Publications}\label{publications}}

Our work is often embedded in publications. We want to make the application of various formatting guidelines as painless as possible. This is one of the main motivations to use markdown and Rmarkdown to document our work. There are more and more conversion tools that automatically convert our longform documentation from markdown or Rmarkdown to the formatting standards of almost any scientific publisher.

\hypertarget{namespace}{%
\section{Namespace}\label{namespace}}

\hypertarget{naming-variables-objects-features}{%
\subsection{Naming variables, objects, features}\label{naming-variables-objects-features}}

In our program code and databases, we are following Open Science naming guidelines. While there are many versions of naming guidelines, we must pick something that works for all and adhere to it.

For ease of programmatic use, we should follow a lowercase, snakecase variable naming.

\begin{verbatim}
## Loading required package: dplyr
\end{verbatim}

\begin{verbatim}
## Warning: package 'dplyr' was built under R version 4.0.4
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'dplyr'
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:stats':
## 
##     filter, lag
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:base':
## 
##     intersect, setdiff, setequal, union
\end{verbatim}

\begin{verbatim}
## Loading required package: snakecase
\end{verbatim}

\begin{tabular}{l|l}
\hline
original & canonical\\
\hline
This Variable & this\_variable\\
\hline
X1 & x\_1\\
\hline
VarName & var\_name\\
\hline
My\_Var & my\_var\\
\hline
that\_variable & that\_variable\\
\hline
\end{tabular}

Variables should describe things starting with general to specific, when possible, or make group selections, such as \texttt{starts\_with()} or \texttt{ends\_with()} easy to use, because \texttt{contains()} can be ambiguous.

\begin{tabular}{l|l}
\hline
original & canonical\\
\hline
spotify\_artist\_id & spotify\_artist\_id\\
\hline
deezer\_id\_artist & deezer\_id\_artist\\
\hline
id\_on\_bandcamp\_track & id\_on\_bandcamp\_track\\
\hline
artist\_bandcamp\_id & artist\_bandcamp\_id\\
\hline
gender\_artist & gender\_artist\\
\hline
\end{tabular}

\begin{itemize}
\tightlist
\item
  \texttt{contains("artist")} selects all artist features
\item
  \texttt{ends\_with("\_id")} selects all ID's. Because \texttt{id} is short, it may appear in words like \texttt{india}. But \texttt{\_id} is unique, particularly if it is the end of a variable name.
\item
  \texttt{starts\_with("spotify")} selects all features related to Spotify, and \texttt{starts\_with("spotify")\ \&\ contains("\_rec\_")} selects all recording features.
\end{itemize}

\hypertarget{metadata}{%
\chapter{Metadata}\label{metadata}}

We follow the metadata definition and concepts of \href{https://mitpress.mit.edu/books/metadata}{Metadata} from the MIT Press Essential Knowledge series.

It is not, Pomerantz tell us, just ``data about data.'' It is a means by which the complexity of an object is represented in a simpler form. For example, the title, the author, and the cover art are metadata about a book. When metadata does its job well, it fades into the background; everyone (except perhaps the NSA) takes it for granted.

Pomerantz explains what metadata is, and why it exists. He distinguishes among different types of metadata---descriptive, administrative, structural, preservation, and use---and examines different users and uses of each type. He discusses the technologies that make modern metadata possible, and he speculates about metadata's future. By the end of the book, readers will see metadata everywhere. Because, Pomerantz warns us, it's metadata's world, and we are just living in it.

\hypertarget{covenant}{%
\chapter{Contributor Covenant Code of Conduct}\label{covenant}}

Placeholder

\hypertarget{our-pledge}{%
\section{Our Pledge}\label{our-pledge}}

\hypertarget{our-standards}{%
\section{Our Standards}\label{our-standards}}

\hypertarget{enforcement-responsibilities}{%
\section{Enforcement Responsibilities}\label{enforcement-responsibilities}}

\hypertarget{scope}{%
\section{Scope}\label{scope}}

\hypertarget{enforcement}{%
\section{Enforcement}\label{enforcement}}

\hypertarget{enforcement-guidelines}{%
\section{Enforcement Guidelines}\label{enforcement-guidelines}}

\hypertarget{correction}{%
\subsection{1. Correction}\label{correction}}

\hypertarget{warning}{%
\subsection{2. Warning}\label{warning}}

\hypertarget{temporary-ban}{%
\subsection{3. Temporary Ban}\label{temporary-ban}}

\hypertarget{permanent-ban}{%
\subsection{4. Permanent Ban}\label{permanent-ban}}

\hypertarget{attribution}{%
\section{Attribution}\label{attribution}}

\hypertarget{appendix}{%
\chapter{Appendix}\label{appendix}}

Placeholder

\hypertarget{markdown-1}{%
\section{Markdown}\label{markdown-1}}

\hypertarget{bookdown}{%
\section{Bookdown}\label{bookdown}}

\hypertarget{tinytex}{%
\section{Tinytex}\label{tinytex}}

  \bibliography{book.bib,packages.bib,antal.bib,ccipolicy.bib,datascience.bib}

\end{document}
